---
title: "Final Project Proposal"
author: "Ashley L. Miller"
date: "2/2/2019"
output:
  html_document:
    highlight: zenburn
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# set knitr options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7.5,
                      fig.height = 4.0)

# load packages
library(rio)
library(here)
library(tidyverse)
library(magrittr)
library(ggforce)

# disable scientific notation
options(scipen = 999)

```

#### Description of the data source (must be publicly available)

```{r load_data}

data <- import(here("data", "DeIntentifiedJML2019Data_Exp1.sav"),
               setclass = "tibble") %>% 
  characterize() %>%
  janitor::clean_names()

```

#### Preliminary ideas (even hand sketches) of different visualizations

  + Identification of the intended audience for each visualization
  + Note, you might consider displaying the same data/relations more than once, with each plot displayed for a different audience.
  + The intended message to be communicated for each plot.
  
```{r preliminary_ideas}



```

